---
title: "EDS 221 Day 9 Interactive Session"
author: "Anna Ramji"
format: html
editor: visual
---

# Day 9 Interactive session

## Setup

```{r}
library(tidyverse)
library(palmerpenguins)
# library(lubridate)
library(kableExtra)
library(janitor)
library(patchwork)
library(reprex)
```

## Warm-up:

Starting from penguins, create a new subset in which you:

1.  Only keep columns species, island, sex, body_mass_g, and flipper_length_mm
2.  Filter to only include Chinstrap and Gentoos
3.  rename the species column to penguin_species
4.  Create a summary table that contains the mean of body mass, the mean of flipper length, and the total count of observations each by species and sex

```{r}
penguins_warmup <- penguins |> 
  select(species, island, sex, body_mass_g, flipper_length_mm) |> 
  filter(species %in% c("Chinstrap", "Gentoo")) |> 
  rename(penguin_species = species) |> 
  group_by(penguin_species, sex) |> 
  summarize(mean_body_mass = mean(body_mass_g, na.rm = TRUE),
            mean_flipper_length = mean(flipper_length_mm, na.rm = TRUE),
            count = n())
```

## **1. Making a reproducible example ({reprex})**

Making a minimum viable example is often the best way to troubleshoot problematic code when you can\'t figure out a solution quickly -- and is *definitely* the best way to share an example of something you\'re struggling with so you\'re most likely to get help. If people can\'t run or play with your code, it\'s much less likely they\'ll be able to offer a solution.

You probably already have {reprex} (part of the tidyverse). Copy code to clipboard and run `reprex()` to make one!

Some guidelines:

-   Ruthlessly simplify

-   Consider using a subset of data (possibly w/ datapasta)

-   Include library calls (e.g. `library(janitor)` in the reprex)

-   Make the minimum viable example of the thing that\'s not working

See more:

-   [**https://community.rstudio.com/t/faq-how-to-do-a-minimal-reproducible-example-reprex-for-beginners/23061**](https://community.rstudio.com/t/faq-how-to-do-a-minimal-reproducible-example-reprex-for-beginners/23061)

-   [**https://reprex.tidyverse.org/articles/reprex-dos-and-donts.html**](https://reprex.tidyverse.org/articles/reprex-dos-and-donts.html)

```{r}

```
